Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-07-27 manuscripts/reply/reply-main.tex
More editing of reply ms.

\section{Introduction}
Biogeographers frequently seek to explain population and species
differentiation on geographical phenomena.
Establishing that a set of population splitting events occurred
at the same time can be a potentially persuasive argument that a set of taxa
were affected by the same geographic events.
% Unfortunately, estimating divergence times precisely when one lacks
% precise information about the rate of molecular evolution is a notoriously
% difficult problem in evolutionary biology.
\citet{Huang2011} introduced an approximate Bayesian approach (implemented in
the software \msb) to estimating the probabilities of models in which multiple
sets of taxa diverge at the same time. 
% This method  does not require a common set of character data that links all
% of the taxa under investigation, but does require the relative mutation
% rates among the taxa and loci to be assumed \emph{a priori}.

@@

\change{
Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice.
}
We then evaluate the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
\citet{Oaks2012}.
In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
made an error by mixing different units of time, which makes the results
presented in their response difficult to interpret (see Supporting Information
for details).
We avoid this error, but still find their approach will often spuriously
support clustered divergence models.
Furthermore, the approach provides another means by which the method can
``escape'' models with large parameter space, which manifests in the preference
of models that exclude from consideration the true values of the model's
parameters.
\change{
Our results, show that it is difficult to choose a uniformly distributed prior
on divergence times that is broad enough to confidently contain the true
parameter values while being narrow enough to avoid spurious support for models
with less parameter space.
}
We also explore the predictions of the potential causes of the bias proposed by
\citet{Oaks2012} and \citet{Hickerson2013}.
We show the primary cause of the method's tendency to incorrectly infer shared
divergences is the disparity in marginal likelihoods across divergence models
imposed by the uniform priors on divergence times.
We also show that sampling error is not a major contributer to the problem.
Fortunately, \citet{Oaks2014dpp} has introduced a method that uses more
flexible probability distributions to accommodate prior uncertainty in
divergence times without inhibiting the marginal likelihoods of models with
more divergence-time parameters, which increases the method's robustness and
power to detect temporal variation in divergences.

@@

Thus, empirical Bayes methods have an uncertain theoretical basis and
do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
empirical Bayesian estimates of the posterior are often too narrow, off-center,
and incorrectly shaped;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
This is not to say that empirical Bayesian approaches are not useful.
Empirical Bayes is a well-studied branch of Bayesian statistics that has given
rise to many inference methods for obtaining parameter estimates that often
exhibit favorable frequentist properties.
% Furthermore, many post-hoc correction methods have been developed for
% estimating confidence-intervals from empirical Bayes estimates of
% distributions that often exhibit well-behaved frequentist coverage
% probabilities
\citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.

@@

In Equation \ref{eq:bayesrule}, we see that the posterior density of any state
$\myTheta{}$ of the model, is the prior density updated by the probability of
the data given $\myTheta{}$ (the likelihood of $\myTheta{}$).
The marginal likelihood of the model only appears as a normalizing constant in
the denominator.
Thus, as long as the prior distribution contains the values of $\myTheta{}$
that maximize the likelihood and the data are strongly informative
relative to the prior, the values of the parameters that maximize the posterior
distribution will be relatively robust to prior choice, even if the posterior
is technically incorrect due to using the data to inform the priors.
However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model that
updates the prior to yield the model's posterior probability.
The integral over the entire parameter space of the likelihood weighted by the
prior density is no longer a normalizing constant, rather it is how the data
inform the posterior probability of the model.
Because the prior probability distributions placed on the model's parameters
have a strong affect on the integrated, or ``average'', likelihood of a model,
Bayesian model choice tends to be much more sensitive to priors than parameter
estimation \citep{Jeffreys1939,Lindley1957}.
Another important difference of Bayesian model choice illustrated by
Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the
posterior probability of a model, is not a function of \myTheta{}, because
it is integrated out of the marginal likelihoods of the candidate models.
Thus, unlike parameter estimates, the estimated posterior probability of a
model is a single value (rather than a distribution) lacking a measure of
posterior uncertainty.

The justification for an empirical Bayesian approach to parameter estimation is
that giving the data more weight relative to the prior (i.e., using the data
twice) will often shift the peak of the estimated distribution nearer to the
true value(s) of the model's parameter(s).
However, there is often no such justification for model selection, because
unlike model parameters, the posterior probabilities of candidate models often
have no clear true values.
% There is no such justification for empirical Bayesian model choice, because
% there are no true values for the model probabilities being estimated.
Model posterior probabilities are inherently measures of our belief in the
models after our prior beliefs are updated by the data being analyzed.
This complicates the meaning of model posterior probabilities when Bayes' rule
is violated by informing priors with the same data to be analyzed.
By using the data twice, we fail to account for prior uncertainty and mislead
our posterior beliefs in the models being compared; we will be over confident
in some models and under confident in others.

@@

\change{
The strong preference for the narrowest priors on divergence times used by
\citet{Hickerson2013} suggests their approach is biased toward models with less
parameter space and, as a consequence, will estimate model-averaged posteriors
dominated by models that exclude true values of the model's parameters.
}
We explored this possibility in two ways.
First, we re-analyzed the Philippines dataset using the model-averaging
approach of \citet{Hickerson2013}, but set one of the prior models with a
uniform prior on divergence times that is unrealistically narrow and almost
certainly excludes most, if not all, of the true divergence times of the 22
taxon pairs.
If small likelihoods of large models cause the method to prefer models with
less parameter space, we expect \msb will preferentially sample from this
incorrect model yielding a posterior that is incorrect (i.e., the
model-averaged posterior will be dominated by an incorrect model that excludes
the truth).
Second, we generated simulated datasets for which the divergence times are
drawn from an exponential distribution and applied the approach of
\citet{Hickerson2013} to each of them to see how often the method excludes the
truth.

@@

Our results show that the model-averaging approach of \citet{Hickerson2013}
does not reduce the method's tendency to support models with less parameter
space.
Rather, the method strongly prefers the prior model with the narrowest
distribution on divergence times across all three of our analyses, even when
this model is incorrect and excludes the true divergence times
of the Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).

However, \citet{Hickerson2013} vetted the priors used in their model-averaging
approach via ``graphical checks,'' in which the summary statistics from 1000
random samples of each prior model are plotted along the first two orthogonal
axes of a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
To determine if such prior predictive analyses would indicate the $M_{1A}$ and
$M_{1B}$ models are problematic, we performed these graphical checks on our
prior models.
Unfortunately, these prior predictive checks provide no warning that these
priors are too narrow (Figure~S\ref{figPCA}).
Rather, the graphs suggest these incorrect priors are ``better fit''
(Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
\citet{Oaks2012} (Figure~S\ref{figPCA}D--F).

@@

\section{Assessing the power of the model-averaging approach of
    \citet{Hickerson2013}}
While our results above clearly demonstrate the risks inherent to the empirical
Bayesian model-choice approach used by \citet{Hickerson2013}, one could justify
such risk if the approach does indeed increase power to detect temporal
variation in divergences and thus decrease bias toward clustered divergence
models.
We assess this possibility using simulation-based analyses.
Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
the 22 population pairs randomly drawn from a uniform distribution, $U(0,
\divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and 2.0, in
\globalcoalunit generations.
All other parameters were identically distributed as the prior models.
As above, we generated a prior sample of $5\e6$ total samples from the five
prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$
(Table~\ref{tabModelChoiceEmpirical}).
For each simulated dataset, we approximated the posterior
by retaining 1000 samples from the prior with the smallest Euclidean distance
from the true summary statistics, as described above.
In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
posterior samples for each of them.

We find that the approach of \citet{Hickerson2013} struggles to estimate the
variance of divergence times (\vmratio{}) across most of the \divt{max}
we simulated, whether evaluating the unadjusted
(Figure~S\ref{figPowerAccuracy}A--F) or GLM-adjusted
(Figure~S\ref{figPowerAccuracy}G--L) posterior estimates.
The method only estimates \vmratio{} relatively well when the simulated
distribution of divergence times is identical to one of the prior models
(Figures~S\ref{figPowerAccuracy}E\&K).
This is consistent with the findings of \citet{Oaks2012}  and
\citet{Hickerson2013} that \msb is highly sensitive to the prior distribution
deviating from the true, underlying distribution of the data.

Furthermore, our results demonstrate that the approach of \citet{Hickerson2013}
consistently infers highly clustered divergences across all the \divt{max} we
simulated (Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
If we follow \citet{Hickerson2013} and use a Bayes factor of greater than 10 as
the criterion for incorrect inference of a single divergence event, we find
that the approach will often strongly support the extreme case of one
divergence event across all our simulation conditions (Figure~\ref{figPower4}E--H \&
S\ref{figPower6}G--L).
Overall, the empirical Bayesian model-averaging approach leads to erroneous
support for highly clustered divergences when populations diverge randomly over
the last $8\globalpopsize$ generations.
To put this on a timescale roughly consistent with a vertebrate mitochondrial
locus, assuming a mutation rate of $1\e{-8}$ per site per generation, this
translates to 10 million generations.
Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
million generations.

% The method is most likely to infer the extreme case of a single divergence
% event shared across all 22 taxa when populations diverged randomly over the
% past \globalcoalunit generations.
% When divergences are random over the past $8\globalpopsize$ generations, the
% most likely inference is only two divergence events, and a single
% divergence is still estimated in more than 10\% of the replicates.
It is interesting to note that as \divt{max} increases, but before the
estimates are finally pulled away from $\numt{} = 1$, the distribution of
\numt{} estimates closely mirror the ``U-shaped'' prior on divergence models
used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
(\citeyear{Oaks2012}) Figure 5B).
This suggests the U-shaped prior coupled with the small marginal likelihoods
of models with many divergence parameters is a major cause of the method's
bias toward clustered divergence models; this is consistent with
\citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.

@@

Also, the results of our power analyses further demonstrate the propensity of
Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
parameter values.
Across all but one of the \divt{max} we simulated, the method favors a model
that excludes the truth in a large proportion of replicates, and across many of
the \divt{max} the preferred model will exclude a large proportion of the true
divergence-time values (Figure~\ref{figPowerExclusion4}A--D \&
S\ref{figPowerExclusion6}A--F).
Importantly, the posterior probability of excluding at least one true
divergence value is also quite high across many of the \divt{max}
(Figure~\ref{figPowerExclusion4}E--H \& S\ref{figPowerExclusion6}G--L).
Only when the data are identically distributed as one of the prior models does
the method avoid excluding the truth more than 5\% of the time
(Figure~\ref{figPowerExclusion4}C).
% Again, this demonstrates the method's sensitivity to priors.

\section{The importance of power analyses to guide applications of \texttt{msBayes}}
\citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
divergence-time prior of 0--1 coalescent units ago.
They found that under these prior conditions \msb can, assuming a
per-site rate of $1.92\e{-8}$ mutations per generation, detect multiple
divergence events among 18 taxa when the true divergences were random over
hundreds of thousands of generations or more.
\citet{Oaks2012} performed similar power analyses under three uniform
divergence-time priors as narrow as 0--5 coalescent units, and found the method
was able to detect multiple events among 22 taxa when divergences were random
over millions of generations.
As recommended by \citet{Oaks2012}, it is important that investigators perform
power analyses to determine the method's power for their dataset, and decide if
\msb has sufficient temporal resolution to address their hypotheses; in the
case of the Philippines dataset, it did not.
When doing so, it is important to consider what prior conditions are
relevant to their empirical system.
A divergence-time prior of 0--5 coalescent units is quite narrow considering
that this expresses the prior belief that all 22 taxon pairs diverged within
this window.
Certainly, it is  rare for there to be enough \emph{a priori} information to
be certain that all taxa diverged within the last 4$\globalpopsize$ generations
(i.e., 0--1 coalescent units).
Also, it seems unlikely that when such prior information is available that
being able to detect multiple divergences on the scale of hundreds of thousands
of generations will add much insight about the evolutionary history of the
taxa.
% ; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), this
% translates to a temporal resolution of 3 million generations or more.

As noted by \citet{Oaks2012}, inferring more than one divergence time
shared across all taxa does not confirm the method is working well when
analyzing data generated under random temporal variation in divergences (e.g.,
an inference of two divergence events could be biogeographically interesting
yet spurious).
Thus, it is important that investigators not limit their assessment of the
method's power to only differentiating inferences of one event or more (i.e.
$\numt{} = 1$ versus $\numt{} > 1$).
Rather, looking at the distribution of estimates, as in \citet{Oaks2012},
provides much more information about the behavior of the method.

\section{The causes of support for models of co-divergence}
To determine how best to improve the behavior of \msb, it is important to
determine the mechanism by which broad uniform priors cause incorrect support
for clustered models of divergence.
It is well established that vague priors can be problematic in Bayesian model
selection.
Models that integrate over more parameter space characterized by low
probability of producing the data and relatively high prior density will have
smaller marginal likelihoods \citep{Jeffreys1939,Lindley1957}.
Given the uniformly distributed priors on divergence times (and other nuisance
parameters) employed in \msb, models with more divergence parameters will be
forced to integrate over \emph{much} greater parameter space, all with equal
prior density, and much of it with low likelihood.
In light of this fundamental statistical issue, it is not surprising that the
method tends to support simple models.

However, \citet{Hickerson2013} conclude that the bias is due to insufficient
sampling.
They argue the widest of the three priors used by \citet{Oaks2012} would
infrequently produce samples with many independent population divergence times
as recent as the estimated gene divergence times presented in \citet{Oaks2012}.
However, this argument assumes the gene divergence times presented in
\citet{Oaks2012} were estimated without error.
These estimates were intended to only provide a rough comparison of the
gene divergence times across the 22 taxa.
These analyses assumed an arbitrary strict mutation rate of $2\e{-8}$ for all
taxa, and are, of course, subject to estimation error.
Furthermore, the branch-length units of the gene trees are in millions of
years, whereas the divergence time prior of \msb is in generations, thus
\citet{Hickerson2013} make the implicit assumption that all 22 Philippine taxa
have a generation time of one year.
The argument of \citet{Hickerson2013} that divergence time estimates of
\citet{Oaks2012} ``should set an upper bound on their prior for \divt{}'' seems
questionable, especially given our findings presented above regarding the
behavior of \msb when empirically informed uniform priors are employed.

@@

This is further demonstrated by the results presented herein that show the
model-averaging approach of \citet{Hickerson2013} prefers models with narrower
\divt{} priors (Table~\ref{tabModelChoiceEmpirical} and
Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerExclusion4})
and models with fewer \divt{} parameters (Figure~\ref{figPower4}).
In all of these analyses, each of the prior models have the same number of
samples.
Thus, while sampling error will always be present in any numerical Bayesian
approximation method, insufficient sampling is an unlikely explanation for the
erroneous support of models with less parameter space.
While analyses that sample each model proportional to their parameter space
could be explored, it is clear that the marginal likelihoods under broad
uniform priors on divergence times will be greater for models with fewer
divergence-time dimensions.

@@

\section{Conclusions}
% We demonstrate how the approximate Bayesian model-choice method implemented in
% \msb can be strongly biased away from models with greater parameter space.
We demonstrate how the approximate Bayesian model-choice method implemented in
\msb can spuriously support models with less parameter space.
As suggested by \citet{Oaks2012}, this is likely caused by the use of uniform
priors on divergence times.
Uniform distributions necessitate the use of priors that place high density in
unlikely regions of parameter space, less the risk of excluding the true
divergence times \emph{a priori}.
These broad uniform priors reduce the marginal likelihoods of models with more
divergence-time parameters.
We show that the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} does not mitigate this bias, but rather causes it to
manifest by sampling predominantly from models that may exclude the true values
of the parameters.
Also, our results show that it is difficult to choose an uniformly distributed
prior on divergence times that is broad enough to confidently contain the true
values of parameters while being narrow enough to avoid bias toward models with
less parameter space.

>>>>

\section{Introduction}
Biogeographers frequently seek to explain population and species
differentiation on geographical phenomena.
Establishing that a set of population splitting events occurred
at the same time can be a potentially persuasive argument that a set of taxa
were affected by the same geographic events.
% Unfortunately, estimating divergence times precisely when one lacks
% precise information about the rate of molecular evolution is a notoriously
% difficult problem in evolutionary biology.
% \citet{Huang2011} introduced an approximate Bayesian approach (implemented in
% the software \msb) to estimating the probabilities of models in which multiple
% sets of taxa diverge at the same time. 
The approximate-Bayesian method, \msb, allows biogeographers to estimate the
probabilities of models in which multiple sets of taxa diverge at the same
time \citep{Hickerson2006,Huang2011}.
% This method  does not require a common set of character data that links all
% of the taxa under investigation, but does require the relative mutation
% rates among the taxa and loci to be assumed \emph{a priori}.

@@

\change{
Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice.
}
We then evaluate the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
\citet{Oaks2012}.
In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
made an error by mixing different units of time, which makes the results
presented in their response difficult to interpret (see Supporting Information
for details).
We avoid this error, but still find their approach will often spuriously
support clustered divergence models.
Furthermore, the approach provides another means by which the method can
``escape'' models with large parameter space, which manifests in the preference
of models that exclude from consideration the true values of the model's
parameters.
\change{
Our results, show that it is difficult to choose a uniformly distributed prior
on divergence times that is broad enough to confidently contain the true
parameter values while being narrow enough to avoid spurious support for models
with less parameter space.
}
We also explore the predictions of the potential causes of the bias proposed by
\citet{Oaks2012} and \citet{Hickerson2013}.
We show the primary cause of the method's tendency to incorrectly infer shared
divergences is the uniform priors on divergence times leading to small marginal
likelihoods of models with more divergence-time parameters.
We also show that sampling error is not a major contributer to the problem.
Fortunately, \citet{Oaks2014dpp} has introduced a method that uses more
flexible probability distributions to accommodate prior uncertainty in
divergence times without overly inhibiting the marginal likelihoods of models
with more divergence-time parameters, which increases the method's robustness
and power to detect temporal variation in divergences.

@@

Thus, empirical Bayes methods have an uncertain theoretical basis and
do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
empirical Bayesian estimates of the posterior are often too narrow, off-center,
and incorrectly shaped;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
This is not to say that empirical Bayesian approaches are not useful.
Empirical Bayes is a well-studied branch of Bayesian statistics that has given
rise to many methods for obtaining parameter estimates that often
exhibit favorable frequentist properties.
% Furthermore, many post-hoc correction methods have been developed for
% estimating confidence-intervals from empirical Bayes estimates of
% distributions that often exhibit well-behaved frequentist coverage
% probabilities
\citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.

@@

In Equation \ref{eq:bayesrule}, we see that the posterior density of any state
$\myTheta{}$ of the model, is the prior density updated by the probability of
the data given $\myTheta{}$ (the likelihood of $\myTheta{}$).
The marginal likelihood of the model only appears as a normalizing constant in
the denominator.
Thus, as long as the prior distribution contains the values of $\myTheta{}$
under which the data are probable and the data are strongly informative
relative to the prior, the values of the parameters that maximize the posterior
distribution will be relatively robust to prior choice, even if the posterior
is technically incorrect due to using the data to inform the priors.
However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model that
updates the prior to yield the model's posterior probability.
The integral over the entire parameter space of the likelihood weighted by the
prior density is no longer a normalizing constant, rather it is how the data
inform the posterior probability of the model.
Because the prior probability distributions placed on the model's parameters
have a strong affect on the integrated, or ``average'', likelihood of a model,
Bayesian model choice tends to be much more sensitive to priors than parameter
estimation \citep{Jeffreys1939,Lindley1957}.
Another important difference of Bayesian model choice illustrated by
Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the posterior
probability of a model, is not a function of \myTheta{}, because the parameters
are integrated out of the marginal likelihoods of the candidate models.
Thus, unlike parameter estimates, the estimated posterior probability of a
model is a single value (rather than a distribution) lacking a measure of
posterior uncertainty.

The justification for an empirical Bayesian approach to parameter estimation is
that giving the data more weight relative to the prior (i.e., using the data
twice) will often shift the peak of the estimated distribution nearer to the
true value(s) of the model's parameter(s).
However, there is no such justification for model selection, because unlike
model parameters, the posterior probabilities of candidate models often have no
clear true values.
% There is no such justification for empirical Bayesian model choice, because
% there are no true values for the model probabilities being estimated.
Model posterior probabilities are inherently measures of our belief in the
models after our prior beliefs are updated by the data being analyzed.
This complicates the meaning of model posterior probabilities when Bayes' rule
is violated by informing priors with the same data to be analyzed.
By using the data twice, we fail to account for prior uncertainty and mislead
our posterior beliefs in the models being compared; we will be over confident
in some models and under confident in others.

@@

\change{
The strong preference for the narrowest prior on divergence times suggests the
approach of \citet{Hickerson2013} is biased toward models with less parameter
space and, as a consequence, will estimate model-averaged posteriors dominated
by models that exclude true values of the model's parameters.
}
We explored this possibility in two ways.
First, we re-analyzed the Philippines dataset using the model-averaging
approach of \citet{Hickerson2013}, but set one of the prior models with a
uniform prior on divergence times that is unrealistically narrow and almost
certainly excludes most, if not all, of the true divergence times of the 22
taxon pairs.
If small likelihoods of large models cause the method to prefer models with
less parameter space, we expect \msb will preferentially sample from this
incorrect model yielding a posterior that is incorrect (i.e., the
model-averaged posterior will be dominated by an incorrect model that excludes
the truth).
Second, we generated simulated datasets for which the divergence times are
drawn from an exponential distribution and applied the approach of
\citet{Hickerson2013} to each of them to see how often the method excludes the
truth.

@@

Our results show that the model-averaging approach of \citet{Hickerson2013}
strongly prefers the prior model with the narrowest distribution on divergence
times across all three of our analyses, even when this model is incorrect and
excludes the true divergence times of the Philippine taxa
(Table~\ref{tabModelChoiceEmpirical}).
However, \citet{Hickerson2013} vetted the priors used in their model-averaging
approach via ``graphical checks,'' in which the summary statistics from 1000
random samples of each prior model are plotted along the first two orthogonal
axes of a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
To determine if such prior predictive analyses would indicate the $M_{1A}$ and
$M_{1B}$ models are problematic, we performed these graphical checks on our
prior models.
Unfortunately, these prior predictive checks provide no warning that these
priors are too narrow (Figure~S\ref{figPCA}).
Rather, the graphs suggest these incorrect priors are ``better fit''
(Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
\citet{Oaks2012} (Figure~S\ref{figPCA}D--F).

@@

\section{Assessing the power of the model-averaging approach of
    \citet{Hickerson2013}}
While our results above clearly demonstrate the risks inherent to the empirical
Bayesian model-choice approach used by \citet{Hickerson2013}, one could justify
such risk if the approach does indeed increase power to detect temporal
variation in divergences and thus decrease bias toward clustered divergence
models.
We assess this possibility using simulations.
Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
the 22 population pairs randomly drawn from a uniform distribution, $U(0,
\divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and 2.0, in
\globalcoalunit generations.
All other parameters were identically distributed as the prior models.
As above, we generated a prior sample of $5\e6$ total samples from the five
prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$
(Table~\ref{tabModelChoiceEmpirical}).
For each simulated dataset, we approximated the posterior
by retaining 1000 samples from the prior with the smallest Euclidean distance
from the true summary statistics, as described above.
In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
posterior samples for each of them.

Our results demonstrate that the approach of \citet{Hickerson2013} consistently
infers highly clustered divergences across all the \divt{max} we simulated
(Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
If we follow \citet{Hickerson2013} and use a Bayes factor of greater than 10 as
the criterion for incorrect inference of a single divergence event, we find
that the approach will often strongly support the extreme case of one
divergence event across all our simulation conditions (Figure~\ref{figPower4}E--H \&
S\ref{figPower6}G--L).
The method also struggles to estimate the variance of divergence times
(\vmratio{}), whether evaluating the unadjusted
(Figure~S\ref{figPowerAccuracy}A--F) or GLM-adjusted
(Figure~S\ref{figPowerAccuracy}G--L) posterior estimates.
Overall, the empirical Bayesian model-averaging approach leads to erroneous
support for highly clustered divergences when populations diverge randomly over
the last $8\globalpopsize$ generations.
To put this on a timescale roughly consistent with a vertebrate mitochondrial
locus (assuming a per-site mutation rate of $1\e{-8}$ per generation) this
translates to 10 million generations.
Assuming a mutation rate consistent with nuclear loci of $1\e{-9}$, this is 100
million generations.

% The method is most likely to infer the extreme case of a single divergence
% event shared across all 22 taxa when populations diverged randomly over the
% past \globalcoalunit generations.
% When divergences are random over the past $8\globalpopsize$ generations, the
% most likely inference is only two divergence events, and a single
% divergence is still estimated in more than 10\% of the replicates.
% It is interesting to note that as \divt{max} increases, but before the
% estimates are finally pulled away from $\numt{} = 1$, the distribution of
% \numt{} estimates closely mirror the ``U-shaped'' prior on divergence models
% used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
% (\citeyear{Oaks2012}) Figure 5B).
% This suggests the U-shaped prior coupled with the small marginal likelihoods
% of models with many divergence parameters is a major cause of the method's
% bias toward clustered divergence models; this is consistent with
% \citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.

@@

Also, the results of our power analyses further demonstrate the propensity of
Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
parameter values.
Across all but one of the \divt{max} we simulated, the method favors a model
that excludes the truth in a large proportion of replicates, and across many of
the \divt{max} the preferred model will exclude a large proportion of the true
divergence-time values (Figure~\ref{figPowerExclusion4}A--D \&
S\ref{figPowerExclusion6}A--F).
Importantly, the posterior probability of excluding at least one true
divergence value is also quite high across many of the \divt{max}
(Figure~\ref{figPowerExclusion4}E--H \& S\ref{figPowerExclusion6}G--L).
% Only when the data are identically distributed as one of the prior models does
% the method avoid excluding the truth more than 5\% of the time
% (Figure~\ref{figPowerExclusion4}C).
% Again, this demonstrates the method's sensitivity to priors.

\section{The importance of power analyses to guide applications of \texttt{msBayes}}
\citet{Hickerson2013} presented a power analysis of \msb under a narrow uniform
divergence-time prior of 0--1 coalescent units ago.
They found that under these prior conditions \msb can, assuming a
per-site rate of $1.92\e{-8}$ mutations per generation, detect multiple
divergence events among 18 taxa when the true divergences were random over
hundreds of thousands of generations or more.
\citet{Oaks2012} performed similar power analyses under three uniform
divergence-time priors as narrow as 0--5 coalescent units, and found the method
was able to detect multiple events among 22 taxa when divergences were random
over millions of generations.
It is important that investigators perform power analyses to determine the
method's power for their dataset, and decide if \msb has sufficient temporal
resolution to address their hypotheses; in the case of the Philippines dataset,
it did not.
When doing so, it is important to consider what prior conditions are
relevant to their empirical system.
A divergence-time prior of 0--5 coalescent units is quite narrow considering
that this expresses the prior belief that all 22 taxon pairs diverged within
this window.
Certainly, it is  rare for there to be enough \emph{a priori} information to
be certain that all taxa diverged within the last 4$\globalpopsize$ generations
(i.e., 0--1 coalescent units).
Also, it seems unlikely that when such prior information is available that
being able to detect more than one divergence event in the face of 18 random
divergences over hundreds of thousands of generations will provide much insight
into the evolutionary history of the taxa.
% ; assuming a rate of mutation consistent with nuclear loci ($1\e{-9}$), this
% translates to a temporal resolution of 3 million generations or more.

Inferring more than one divergence time shared across all taxa does not confirm
the method is working well when analyzing data generated under random temporal
variation in divergences (e.g., an inference of two divergence events could be
biogeographically interesting yet spurious).
Thus, it is important that investigators not limit their assessment of the
method's power to only differentiating inferences of one event or more (i.e.
$\numt{} = 1$ versus $\numt{} > 1$).
Rather, looking at the distribution of estimates, as in Figure~\ref{figPower4}
and \citet{Oaks2012}, provides much more information about the behavior of the
method.

\section{The causes of support for models of co-divergence}
To determine how best to improve the behavior of \msb, it is important to
determine the mechanism by which broad uniform priors cause incorrect support
for clustered models of divergence.
It is well established that vague priors can be problematic in Bayesian model
selection.
Models that integrate over more parameter space characterized by low
probability of producing the data and relatively high prior density will have
smaller marginal likelihoods \citep{Jeffreys1939,Lindley1957}.
Given the uniformly distributed priors on divergence times employed in \msb,
models with more divergence parameters will be ``averaged'' over much greater
parameter space, all with equal prior weight, and much of it with low
likelihood.
In light of this fundamental statistical issue, it is not surprising that the
method tends to support simple models.

However, \citet{Hickerson2013} conclude that the bias is due to insufficient
sampling.
They argue the widest of the three priors used by \citet{Oaks2012} would
infrequently produce samples with many independent population divergence times
as recent as the estimated gene divergence times presented in \citet{Oaks2012}.
However, this argument assumes the gene divergence times presented in
\citet{Oaks2012} were estimated without error.
These estimates were intended to only provide a rough comparison of the
gene divergence times across the 22 taxa.
These analyses assumed an arbitrary strict per-site mutation rate of $2\e{-8}$
per generation for all taxa, and are, of course, subject to estimation error.
Furthermore, the branch-length units of the gene trees are in millions of
years, whereas the divergence time prior of \msb is in generations, thus
\citet{Hickerson2013} make the implicit assumption that all 22 Philippine taxa
have a generation time of one year.
The argument of \citet{Hickerson2013} that divergence-time estimates of
\citet{Oaks2012} should set an upper bound on the prior for divergence times
seems questionable, especially given our findings presented above regarding the
behavior of \msb when empirically informed uniform priors are employed.

@@

This is further demonstrated by the results presented herein that show the
model-averaging approach of \citet{Hickerson2013} prefers models with narrower
\divt{} priors (Table~\ref{tabModelChoiceEmpirical} and
Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerExclusion4})
and models with fewer \divt{} parameters (Figure~\ref{figPower4}).
In all of these analyses, each of the prior models have the same number of
samples.
Thus, while sampling error will always be present in any numerical Bayesian
approximation method, insufficient sampling is an unlikely explanation for the
erroneous support for models with less parameter space.
While analyses that sample each model proportional to their parameter space
could be explored, it is clear that the marginal likelihoods under broad
uniform priors on divergence times will be greater for models with fewer
divergence-time dimensions.

@@

\section{Conclusions}
% We demonstrate how the approximate Bayesian model-choice method implemented in
% \msb can be strongly biased away from models with greater parameter space.
We demonstrate how the approximate Bayesian model-choice method implemented in
\msb can spuriously support models with less parameter space.
This is caused by the use of uniform priors on divergence times.
Uniform distributions necessitate the use of priors that place high density in
unlikely regions of parameter space, less the risk of excluding the true
divergence times \emph{a priori}.
These broad uniform priors reduce the marginal likelihoods of models with more
divergence-time parameters.
We show that the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} does not mitigate this bias, but rather causes it to
manifest by sampling predominantly from models that may exclude the true values
of the parameters.
Also, our results show that it is difficult to choose an uniformly distributed
prior on divergence times that is broad enough to confidently contain the true
values of parameters while being narrow enough to avoid erroneous support for
models with less parameter space.
